# descriptive


How does Gifted and Talented affect student test scores?

## Data Overview

### Data cleaning

Using OpenData NYC, we pull aggregate K-8 attendance records and socio-economic variables for all 1125 public schools in NYC.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#install and load packages
library(tidyverse)
library(readr)
library(tidyverse)
library(kableExtra)


merged<- read_csv("data/merged.csv")
head(merged[1:3,c(1:6,8:9)]) %>%
  kable(caption = "Sample Data Frame",
        col.names = c(" ", "DBN", "School Name",
                      "ENI", "% Male", "% Black", "%SWD", "% Poverty"))
```


### Visualizing the Treatment (GT option)

Treatment in each borough. 101 schools have the gifted option, 653 no gifted option


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
school_wide <- read_csv("data/merged.csv")
ela_wide <- read_csv("raw/ela_wide (1).csv")
math_wide <- read_csv("raw/math_wide (1).csv")
## bring together
data <- merge(ela_wide, math_wide, by = "DBN")
data <- merge(data, school_wide, by = "DBN")
## small table
small <- data %>%
  select(DBN, `School Name`, class_option, `3 2019.y`, `4 2019.y`, `5 2019.y`,
         `3 2019.x`, `4 2019.x`, `5 2019.x`, `Total Enrollment`, `% Poverty`,`% Students with Disabilities`, `Economic Need Index`, borough, school_type) %>%
  mutate(GT_option = case_when(class_option %in% c("GT", "SC and GT") ~ "gifted option",
                               class_option %in% c("no option", "SC") ~ "no gifted option"),
         SC_option = case_when(class_option %in% c("SC", "SC and GT") ~ "SC option",
                               class_option %in% c("no option", "GT") ~ "no SC option"),
         treatment = ifelse(GT_option == "gifted option", 1, 0))

#### 
### STEP 2: Prevalence of the Option, Likelihood
####
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}
liklihood <- small %>% 
  group_by(borough) %>%
  mutate(total=n()) %>%
  group_by(class_option, borough) %>%
  summarise(n=n(),
            ratio = n/total) %>%
  mutate(percentage = percent(ratio)) %>%
  distinct() %>%
  select(class_option, borough, percentage, ratio, n) %>%
  arrange(borough)
liklihood$ratio2 <- percent(liklihood$ratio)
library(viridis)
ggplot(data = liklihood, aes(fill = class_option, y = ratio,
                       x = borough , label = ratio2))+
  geom_bar(position = "dodge", stat = "identity") +
  theme(
    legend.position = "none",
    plot.title = element_text(size=11),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    panel.background = element_rect(fill = 'white'),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  scale_y_continuous(labels = scales::percent)+
  geom_text(size = 4, position=position_dodge(width=1), vjust=-1, hjust = .4)

```

### Raw Differences in Outcome and Covariates 

```{r}

small$`3 2019.ela` <-gsub(",","",small$`3 2019.y`,fixed = TRUE)
small$`3 2019.math` <-gsub(",","",small$`3 2019.x`,fixed = TRUE)
library(tidyverse)
small<- small %>%
  group_by(treatment) %>%
  na.omit() %>%
  filter(`3 2019.ela` != "s") 
small %>%
  summarise(mean_third_ela = mean(as.numeric(`3 2019.ela`)),
            mean_third_math = mean(as.numeric(`3 2019.math`))) %>%
   kable(caption = "Outcome Differences by Treatment",
        col.names= c("Treatment", "Mean ELA", "Mean Math"),
        digits = c(0,2,2))
# 
# table(small$GT_option)
### GT Option by Economic Need, Poverty, Enrollment
small%>%
  group_by(GT_option) %>%
  summarise(mean_ec = mean(`Economic Need Index`),
            mean_poverty = mean(`% Poverty`),
            total = mean(`Total Enrollment`))
small%>%
  group_by(SC_option) %>%
  summarise(mean_ec = mean(`Economic Need Index`),
            mean_poverty = mean(`% Poverty`),
            total = mean(`Total Enrollment`))

#### GT Option by Borough
### boroughs
small %>%
  group_by(borough, GT_option) %>%
  summarise(mean_ec = mean(`Economic Need Index`),
            mean_poverty = mean(`% Poverty`),
            total = mean(`Total Enrollment`))


### GT Option by School Type *** NOTE: Middle Schools again have the biggest difference
small %>%
  group_by(school_type, GT_option) %>%
  summarise(mean_ec = mean(`Economic Need Index`),
            mean_poverty = mean(`% Poverty`),
            total = mean(`Total Enrollment`))
```


## Naive regression models (without matching)

Treatment (GT) higher ELA and Math scores

```{r}

mod <- lm(`3 2019.ela` ~ treatment + `Economic Need Index` + `Total Enrollment`, data =small)
summary(mod)
table(small$treatment)
mod <- lm(`3 2019.math` ~ treatment, data =small)
summary(mod)

```



