# Simulation and Loops

### ATE results Loop

Do the ATE results hold for different DVs, across grade levels and years?

```{r message=FALSE, warning=FALSE, include=FALSE}
#install and load packages
library(tidyverse)
library(readr)
library(tidyverse)
library(kableExtra)

data<- readRDS("/cloud/project/data/total_data_wide.RDS")
head(data[1:3,c(1:6,8:9)]) %>%
  kable(caption = "Sample Data Frame",
        col.names = c(" ", "DBN", "School Name",
                      "ENI", "% Male", "% Black", "%SWD", "% Poverty"))

## rename data frame
data <- data %>%
  rename(TotalEnrollment = `Total Enrollment`,
         ENI = `Economic Need Index`,
         PercentBlack = `% Black`,
         PercentWhite = `% White`,
         PercentSWD = `% Students with Disabilities`,
         PercentPoverty = `% Poverty`,
         PercentAsian = `% Asian`,
         PercentHispanic = `% Hispanic`,
         PercentELL = `% English Language Learners`,
         Percent_Attendance = `% Attendance`,
         Percent_Chron_Absent = `% Chronically Absent`,
         PercentMale= `% Male`)
data <- data %>%
  mutate(GT_option = case_when(class_option %in% c("GT", "SC and GT") ~ "gifted option",
                               class_option %in% c("no option", "SC") ~ "no gifted option"),
         SC_option = case_when(class_option %in% c("SC", "SC and GT") ~ "SC option",
                               class_option %in% c("no option", "GT") ~ "no SC option"),
         treatment = ifelse(GT_option == "gifted option", 1, 0))


```


```{r message=FALSE, warning=FALSE}

grade_vector <- c("3", "4", "5", "6")
year_vector <- c('2013.y', '2014.y', '2015.y', '2016.y', '2017.y', '2018.y', '2019.y',
                '2013.x', '2014.x', '2015.x', '2016.x', '2017.x', '2018.x', '2019.x')

empty_list <- list()
year_list<- list()
for (i in 1:length(grade_vector)) {
grade <- grade_vector[i]
df<- data.frame("year" = NA, "grade" = NA, "unmatched_regression_estimate" = NA, "p value" = NA)
for (b in 1:length(year_vector)){
year <- year_vector[b]
gy <- paste(grade, year)
data$outcome <-gsub(",","",data[,as.character(gy)],fixed = TRUE)
data <- data %>% filter(!is.na(outcome))
mod1 <- lm(as.numeric(outcome) ~ as.factor(treatment), data = data)
df[b,1]<- year
df[b,2]<- grade
df[b,3] <- mod1$coefficients[[2]]
df[b,4] <- summary(mod1)$coefficients[,4][[2]]
empty_list[[i]] <- df
}
}
dv_results <- as.data.frame(do.call(rbind, empty_list))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

df <- dv_results %>%
  separate(year, c("Year", "Subject")) %>%
  mutate(Subject = recode(Subject,
                          "y" = "English",
                          "x" = "Math")) %>%
  mutate(sig = ifelse(p.value < .001, "*", " "))

df %>%
  filter(Subject == "English") %>%
  ggplot(aes(x= as.numeric(Year), y= as.numeric(unmatched_regression_estimate), 
            group = as.factor(grade), color = as.factor(grade))) +
  geom_point(aes(shape = sig, size = 3)) +
  scale_shape_manual(values = c(1,8))+
  geom_line() +
  ylab("Estimate of English Score Increase")+
  xlab("School Year") +
  theme_minimal() +
  ggtitle("Matched Estimates of GT on English Scores \n for Grades 3-6 over Years 2014-2018") +
  guides(color = guide_legend("Grade")) +
  labs(caption = "Matched Estimates on 5 Covariates, sig tested at <.001 level \n
       * denotes significant effects") +
  theme(axis.title = element_text(size = 18),
        axis.text.x = element_text(size = 12, vjust =6),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 18)) +
  scale_x_discrete(limits = c(2013, 2014, 2015, 2016, 2017, 2018, 2019)) 

df %>%
  filter(Subject == "Math") %>%
  ggplot(aes(x= as.numeric(Year), y= as.numeric(unmatched_regression_estimate), 
            group = as.factor(grade), color = as.factor(grade))) +
  geom_point(aes(shape = sig, size = 3)) +
  scale_shape_manual(values = c(1,8))+
  geom_line() +
  ylab("Estimate of Math Score Increase")+
  xlab("School Year") +
  theme_minimal() +
  ggtitle("Matched Estimates of GT on Math Scores \n for Grades 3-6 over Years 2014-2018") +
  guides(color = guide_legend("Grade")) +
  labs(caption = "Matched Estimates on 5 Covariates, sig tested at <.001 level \n
       * denotes significant effects") +
  theme(axis.title = element_text(size = 18),
        axis.text.x = element_text(size = 12, vjust =6),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 18)) +
  scale_x_discrete(limits = c(2013, 2014, 2015, 2016, 2017, 2018, 2019)) 

```




### Group structured data (when the data has a multilevel structure)




### Estimators

Three categories of estimators: estimators that model sample selection (propensity scores), estimators that model the outcome directly, and estimators that model both the sample selection and the outcome

* IPW
* BART
* Outcome Modeling 
* Weight Trimming
* Doubly Robust Methods

* Interaction of weights with multi-level matching... how bias/precision is affected
* Generalizability to different DVs (quant optimal/policy relevant)









